\section{Game Rules and Formalization}
\label{sec:rules}

\subsection{Setup}

German Whist is played by two players with a standard 52-card deck. Cards are
ranked Ace (high), King, Queen, Jack, 10, 9, 8, 7, 6, 5, 4, 3, 2 (low) within
each suit. Each player is dealt 13 cards, and the remaining 26 cards form the
\emph{stock}. The top card of the stock is placed face-up; its suit determines
the \emph{trump suit} for the entire game.

\subsection{Trick-Taking Rules}

The non-dealer leads to the first trick. Standard trick-taking rules apply:
\begin{itemize}
  \item The leader may play any card from their hand.
  \item The follower \emph{must follow suit} if able (play a card of the same
    suit as the lead). If unable, they may play any card.
  \item If both cards are the same suit, the higher-ranked card wins.
  \item If the follower plays a different suit, the leader wins
    \emph{unless} the follower played a trump, in which case the trump wins.
\end{itemize}

\subsection{Phase 1: Card Acquisition (Tricks 1--13)}

After each trick in Phase~1:
\begin{enumerate}
  \item The trick winner takes the face-up card from the stock and adds it to
    their hand.
  \item The trick loser takes the next (face-down) card from the stock,
    without revealing it.
  \item The next card on the stock is turned face-up.
  \item The trick winner leads the next trick.
\end{enumerate}

Crucially, Phase~1 tricks do not count toward the final score. The sole purpose
of Phase~1 is to \emph{build a strong hand} for Phase~2. Since both players
draw a card after each trick, hand sizes remain at 13 throughout Phase~1.

\subsection{Phase 2: Trick-Taking (Tricks 14--26)}

After the stock is exhausted (13 tricks), Phase~2 begins. Play continues
normally but without card drawing. The winner of the 13th trick in Phase~1
leads first in Phase~2. Whoever wins the \emph{majority} of the 13 Phase~2
tricks wins the game. Since 13 is odd, draws are impossible.

\subsection{Formal State Space}

We define the game state as a tuple:
\[
  s = (H_0, H_1, S, \tau, \ell, \phi, c_\ell, t_0, t_1)
\]
where $H_0, H_1$ are the players' hands (sets of cards), $S$ is the stock
(ordered sequence), $\tau$ is the trump suit, $\ell \in \{0,1\}$ is the
current leader, $\phi \in \{1,2\}$ is the phase, $c_\ell$ is the card led
(if any), and $t_0, t_1$ are Phase~2 trick counts.

\subsubsection{Information Sets}

In Phase~1, each player's \emph{information set} consists of:
\begin{itemize}
  \item Their own hand
  \item All cards played in previous tricks
  \item All face-up cards observed (including those taken by either player)
  \item The current face-up card on the stock
\end{itemize}
The opponent's hand composition and the order of remaining stock cards are
unknown, making Phase~1 a game of \emph{imperfect information}.

In Phase~2, a player who has perfectly tracked all cards can deduce the
opponent's exact hand, making Phase~2 a game of \emph{perfect information}.
