\section{Phase 2 Analysis}
\label{sec:phase2}

Phase~2 is a 13-trick perfect-information game. With exact minimax, we can
analyze optimal play patterns.

\subsection{Endgame Properties}

Several structural properties of the Phase~2 endgame are notable:

\begin{enumerate}
  \item \textbf{No draws:} With 13 tricks and the majority needed to win,
    every game has a winner (7+ tricks).
  \item \textbf{Lead advantage:} The player on lead has an inherent advantage,
    as they can dictate the suit of play. Our simulations show the Phase~2
    leader wins approximately 55--60\% of the time in random positions.
  \item \textbf{Trump control:} Holding more trumps than the opponent is the
    single strongest predictor of Phase~2 success, as trumps both win tricks
    directly and prevent the opponent from cashing side-suit winners.
\end{enumerate}

\subsection{Key Strategic Principles}

\subsubsection{Cash Winners First}

When on lead with sure winners (top cards in a suit), play them immediately.
Delay risks the opponent gaining the lead and establishing their own suits.

\subsubsection{Long Suits}

A long suit (e.g., 5+ cards) becomes increasingly powerful as the opponent
runs out of that suit. After the opponent is void, remaining cards in the
suit are all winners---provided we can retain the lead to play them.

\subsubsection{Trump Management}

Trumps serve dual purposes: winning tricks directly when led, and
\emph{ruffing} (trumping the opponent's side-suit leads). Optimal play
balances these uses:
\begin{itemize}
  \item Drawing trumps when ahead in the trump suit removes the opponent's
    ability to ruff.
  \item Preserving trumps when behind in a side suit enables defensive
    ruffing.
\end{itemize}

\subsection{Minimax Performance}

Our hybrid solver (3-trick heuristic lookahead + exact solve at $\leq 10$
cards) achieves:
\begin{itemize}
  \item Median solve time: $< 100$ms per move
  \item Maximum solve time: $< 500$ms per move
  \item Exact play for positions with $\leq 10$ cards per player
  \item Near-optimal play for the opening Phase~2 moves via quick-trick
    heuristic evaluation
\end{itemize}

The exact solve threshold is set high enough that the majority of Phase~2
decisions---where suit depletion, trump battles, and endgame squeeze plays
occur---are resolved optimally.
