\section{Simulation Results}
\label{sec:results}

We conducted extensive simulations pitting different player types against each
other to quantify the effectiveness of various strategies.

\subsection{Player Types}

\begin{itemize}
  \item \textbf{Random:} Selects uniformly at random from legal moves.
  \item \textbf{Heuristic:} Rule-based player using hand-coded strategy for
    both phases: win high face-up cards in Phase~1, lead aces and long suits
    in Phase~2.
  \item \textbf{AI:} Rule-based heuristic for Phase~1, alpha-beta minimax
    (exact at $\leq 10$ cards) for Phase~2, with full card counting.
\end{itemize}

\subsection{Head-to-Head Results}

\begin{center}
\begin{tabular}{llccc}
\toprule
Player 0 & Player 1 & Games & P0 Win\% & Time/Game \\
\midrule
Random & Random & 1000 & 52.0\% & $<0.01$s \\
Heuristic & Random & 1000 & 93.2\% & $<0.01$s \\
AI & Random & 500 & 91.4\% & $<0.1$s \\
AI & Heuristic & 500 & 52.6\% & $<0.1$s \\
Heuristic A & Heuristic B & 1000 & 47.4\% & $<0.01$s \\
\bottomrule
\end{tabular}
\end{center}


\subsection{Key Findings}

\subsubsection{Heuristic vs.\ Random}

The heuristic player dominates random play, winning over 91\% of games with an
average trick margin of 2.7 tricks in Phase~2. This demonstrates that even
simple strategic reasoning---winning good face-up cards, leading aces---provides
an enormous advantage.

\subsubsection{AI vs.\ Random}

The AI wins 91.4\% of games against random play, closely matching the
heuristic's 93.2\%. Both use the same rule-based Phase~1 strategy; the AI's
slight disadvantage comes from its heuristic opening in Phase~2 (for hands
$>10$ cards) before switching to exact minimax.

\subsubsection{AI vs.\ Heuristic}

Against the heuristic opponent, the AI wins 52.6\% of 500 games. The AI's edge
comes from its Phase~2 minimax solver, which plays exactly when hands reach 10
or fewer cards. The card-counting ability ensures the AI always knows the
opponent's exact hand entering Phase~2. However, with a 95\% CI of
[48.2\%, 57.0\%], this result is not statistically significant---a larger
sample would be needed to confirm the advantage. The direction is consistent
with expectations, since exact minimax should outperform heuristic play in
Phase~2, but the effect size is small relative to the variance inherent in
Phase~1 card acquisition.

\subsubsection{First-Player Advantage}

Analysis of random-vs-random games shows the non-dealer (Player~0, who leads
first) wins 52.0\% of games (95\% CI: [48.9\%, 55.1\%])---consistent with no
significant first-player advantage. Interestingly, in heuristic-vs-heuristic
games, the \emph{dealer} (Player~1) wins slightly more often at 52.6\%,
suggesting a slight dealer advantage in skilled play, possibly because the
dealer gets the last draw from the stock in Phase~1.

\subsection{Score Distributions}

The distribution of Phase~2 trick differentials reveals the variance in game
outcomes:
\begin{itemize}
  \item Random vs.\ Random: approximately normal distribution centered at 0,
    with standard deviation of $\approx 3.0$ tricks.
  \item Heuristic vs.\ Random: heavily skewed toward the heuristic, with a
    mode at $+3$ to $+5$ tricks.
  \item AI vs.\ Heuristic: moderately skewed toward the AI, with occasional
    large wins in both directions reflecting the variance in Phase~1 card
    acquisition.
\end{itemize}

\subsection{Phase 1 Impact}

A critical finding from our simulations is the importance of Phase~1 play on
game outcomes. Players who acquire more trump cards during Phase~1 win
Phase~2 significantly more often. Specifically:
\begin{itemize}
  \item Holding 4+ trumps entering Phase~2 correlates with a win rate above
    65\%.
  \item Holding 6+ trumps entering Phase~2 correlates with a win rate above
    80\%.
  \item Void suits (enabling ruffing) combined with trump length is the
    strongest hand configuration.
\end{itemize}

\subsection{Computational Performance}

\begin{center}
\begin{tabular}{lccc}
\toprule
Component & Median Time & Max Time & Notes \\
\midrule
Phase 2 minimax ($\leq 10$ cards) & 30ms & 100ms & Exact solve \\
Phase 1 heuristic & $<1$ms & $<1$ms & Rule-based \\
Full AI game & $<100$ms & $<200$ms & 26 moves per game \\
Heuristic game & $<10$ms & $<10$ms & Instant \\
Random game & $<1$ms & $<1$ms & Instant \\
\bottomrule
\end{tabular}
\end{center}
